---
title: "String Manipulation With stringr"
author: "Dr. Hua Zhou"
date: "Feb 6, 2018"
subtitle: Biostat M280
output:
  ioslides_presentation:
    smaller: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 5, fig.height = 3.5, fig.align = 'center',
  cache = FALSE)
```

# String | r4ds chapter 14

## stringr

- stringr pacakge, by Hadley Wickham, provides utilities for handling strings.

- Included in tidyverse.
    ```{r}
    library("tidyverse")
    ```

## Basics

- Strings are enclosed by double quotes orx single quotes:
    ```{r}
    string1 <- "This is a string"
    string2 <- 'If I want to include a "quote" inside a string, I use single quotes'
    ```

- Literal single or double quote:
    ```{r}
    double_quote <- "\"" # or '"'
    single_quote <- '\'' # or "'"
    ```

----

- Printed representation:
    ```{r}
    x <- c("\"", "\\")
    x
    ```
vs `writeLines()`:
    ```{r}
    writeLines(x)
    ```
    
----    

- Other special characters: `"\n"` (new line), `"\t"` (tab), ... Check
    ```{r, eval = FALSE}
    ?"'"
    ```
for a complete list.

- Unicode
    ```{r}
    x <- "\u00b5"
    x
    ```

- Character vector (vector of strings):
    ```{r}
    c("one", "two", "three")
    ```

## String length

- Length of a single string:
    ```{r}
    str_length("R for data science")
    ```

- Lengths of a character vector:
    ```{r}
    str_length(c("a", "R for data science", NA))
    ```

## Combining strings

- Combine two or more strings
    ```{r}
    str_c("x", "y")
    str_c("x", "y", "z")
    ```

- Separator:
    ```{r}
    str_c("x", "y", sep = ", ")
    ```

----

- `str_c()` is vectorised:
    ```{r}
    str_c("prefix-", c("a", "b", "c"), "-suffix")
    ```

- Objects of length 0 are silently dropped:
    ```{r}
    name <- "Hadley"
    time_of_day <- "morning"
    birthday <- FALSE
    
    str_c(
      "Good ", time_of_day, " ", name,
      if (birthday) " and HAPPY BIRTHDAY",
      "."
    )
    ```

----

- Combine a vector of strings:
    ```{r}
    str_c(c("x", "y", "z"))
    str_c(c("x", "y", "z"), collapse = ", ")
    ```
    
## Subsetting strings

- By position:
    ```{r}
    str_sub("Apple", 1, 3)
    x <- c("Apple", "Banana", "Pear")
    str_sub(x, 1, 3)
    ```

- Negative numbers count backwards from end:
    ```{r}
    str_sub(x, -3, -1)
    ```

----

- Out of range:
    ```{r}
    str_sub("a", 1, 5)
    str_sub("a", 2, 5)
    ```

- Assignment to a substring:
    ```{r}
    str_sub(x, 1, 1) <- str_to_lower(str_sub(x, 1, 1))
    x
    ```

## Matching patterns with regular expressions

- `str_view()` shows the first match;  
`str_view_all()` shows all matches.

- Match exact strings:
    ```{r}
    x <- c("apple", "banana", "pear")
    str_view(x, "an")
    str_view_all(x, "an")
    ```

----

- `.` matches any character apart from a newline:
    ```{r}
    str_view(x, ".a.")
    ```

- To match a literal `.`:
    ```{r}
    str_view(c("abc", "a.c", "bef"), "a\\.c")
    ```

- To match a literal `\`:
    ```{r}
    str_view("a\\b", "\\\\")
    ```

## Anchors

- `^` matches the start of the string:
    ```{r}
    x <- c("apple", "banana", "pear")
    str_view(x, "^a")
    ```

- `$` matches the end of the string:
    ```{r}
    str_view(x, "a$")
    ```

----

- To force a regular expression to only match a complete string:
    ```{r}
    x <- c("apple pie", "apple", "apple cake")
    str_view(x, "^apple$")
    ```

----

- Other special matches:

    - `\d`: matches any digit.
    
    - `\s`: matches any whitespace (e.g. space, tab, newline).
    
    - `[abc]`: matches a, b, or c.
    
    - `[^abc]`: matches anything except a, b, or c.
    
    - alternation
    ```{r}
    str_view(c("grey", "gray"), "gr(e|a)y")
    ```
    ```{r}
    str_view(c("grey", "gray"), "gr[ea]y")
    ```
    
## Repetition

- 
`?`: 0 or 1  
`+`: 1 or more  
`*`: 0 or more

- 
    ```{r}
    x <- "1888 is the longest year in Roman numerals: MDCCCLXXXVIII"
    str_view(x, "CC?")
    # greedy matches
    str_view(x, "CC+")
    # greedy matches
    str_view(x, 'C[LX]+')
    ```
    
----

- Specify number of matches:  
`{n}`: exactly n   
`{n,}`: n or more  
`{,m}`: at most m  
`{n,m}`: between n and m

-
    ```{r}
    str_view(x, "C{2}")
    # greedy matches
    str_view(x, "C{2,}")
    # greedy matches
    str_view(x, "C{2,3}")
    ```

----

- Greedy (default) vs lazy (put `?` after repetition):
    ```{r}
    # lazy matches
    str_view(x, 'C{2,3}?')
    # lazy matches
    str_view(x, 'C[LX]+?')
    ```

## Grouping and backreference

- Parentheses define groups, which can be back-referenced as `\1`, `\2`, ...
    ```{r}
    str_view(fruit, "(..)\\1", match = TRUE)
    ```
    
## Detect matches

- 
    ```{r}
    x <- c("apple", "banana", "pear")
    str_detect(x, "e")
    ```

- Vector `words` contains 1000 commonly used words:
    ```{r}
    length(words)
    head(words)
    ```

----

- 
    ```{r}
    # How many common words start with t?
    sum(str_detect(words, "^t"))
    # What proportion of common words end with a vowel?
    mean(str_detect(words, "[aeiou]$"))
    ```
    
----

- Find workds that end with `x`:
    ```{r}
    words[str_detect(words, "x$")]
    ```
same as
    ```{r}
    str_subset(words, "x$")
    ```
    
----

- Filter a data frame:
    ```{r}
    df <- tibble(
      word = words, 
      i = seq_along(word)
    )
    df %>% 
      filter(str_detect(words, "x$"))
    ```

----

- `str_count()` tells how many matches are found:
    ```{r}
    x <- c("apple", "banana", "pear")
    str_count(x, "a")
    ```
    ```{r}
    # On average, how many vowels per word?
    mean(str_count(words, "[aeiou]"))
    ```

- Matches never overlap:
    ```{r}
    str_count("abababa", "aba")
    str_view_all("abababa", "aba")
    ```

----

- Mutate a data frame:
    ```{r}
    df %>% 
      mutate(
        vowels = str_count(word, "[aeiou]"),
        consonants = str_count(word, "[^aeiou]")
      )
    ```

## Extract matches

- `sentences` is a collection of 720 phrases:
    ```{r}
    length(sentences)
    head(sentences)
    ```
    
- Suppose we want to find all sentences that contain a colour.

----

- Create a collection of colours:
    ```{r}
    colours <- c("red", "orange", "yellow", "green", "blue", "purple")
    colour_match <- str_c(colours, collapse = "|")
    colour_match
    ```

- Select the sentences that contain a colour, and then extract the colour to figure out which one it is:    
    ```{r}
    has_colour <- str_subset(sentences, colour_match)
    matches <- str_extract(has_colour, colour_match)
    head(matches)
    ```

----

- `str_extract()` only extracts the first match.

-
    ```{r}
    more <- sentences[str_count(sentences, colour_match) > 1]
    str_view_all(more, colour_match)
    ```

- `str_extract_all()` extracts all matches:
    ```{r}
    str_extract_all(more, colour_match)
    ```

----

- Setting `simplify = TRUE` in `str_extract_all()` will return a matrix with short matches expanded to the same length as the longest:
    ```{r}
    str_extract_all(more, colour_match, simplify = TRUE)
    ```
    ```{r}
    x <- c("a", "a b", "a b c")
    str_extract_all(x, "[a-z]", simplify = TRUE)
    ```

## Grouped matches

- `str_extract()` gives us the complete match:
    ```{r}
    noun <- "(a|the) ([^ ]+)"

    has_noun <- sentences %>%
      str_subset(noun) %>%
      head(10)
    has_noun %>%
      str_extract(noun)
    ```

----

- `str_match()` gives each individual component:
    ```{r}
    has_noun %>%
      str_match(noun)
    ```

----

- `tidyr::extract()` works with tibble:
    ```{r}
    tibble(sentence = sentences) %>%
      tidyr::extract(
        sentence, c("article", "noun"), "(a|the) ([^ ]+)",
        remove = FALSE
      )
    ```
    
## Replacing matches

- Replace the first match:
    ```{r}
    x <- c("apple", "pear", "banana")
    str_replace(x, "[aeiou]", "-")
    ```

- Replace all matches:
    ```{r}
    str_replace_all(x, "[aeiou]", "-")
    ```

- Multiple replacement:
    ```{r}
    x <- c("1 house", "2 cars", "3 people")
    str_replace_all(x, c("1" = "one", "2" = "two", "3" = "three"))
    ```

----

- Back-reference:
    ```{r}
    # flip the order of the second and third words
    sentences %>% 
      str_replace("([^ ]+) ([^ ]+) ([^ ]+)", "\\1 \\3 \\2") %>% 
      head(5)
    ```

## Splitting

- Split a string up into pieces:
    ```{r}
    sentences %>%
      head(5) %>% 
      str_split(" ")
    ```

----
    
- Use `simplify = TRUE` to return a matrix:    
    ```{r}
    sentences %>%
      head(5) %>% 
      str_split(" ", simplify = TRUE)
    ```
    
