---
title: "Biostat M280 Homework 3"
subtitle: Due ??? @ 11:59PM
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = FALSE)
```

```{r}
sessionInfo()
```

# LA Parking Ticket

## CSV file

The `/home/m280-data/la_parking/LA_Parking_Citations.csv` file contains information about parking tickets in LA City. It is downloaded from [LA Open Data Portal](https://data.lacity.org/A-Well-Run-City/Parking-Citations/wjz9-h9np). 
```{bash, eval = (Sys.info()[["sysname"]] == "Linux")}
ln -sf /home/m280-data/la_parking/LA_Parking_Citations.csv LA_Parking_Citations.csv
```

The CSV file is about 500MB.
```{bash}
ls -l LA_Parking_Citations.csv
```

How many rows?
```{bash}
wc -l < LA_Parking_Citations.csv
```

First few lines:
```{bash}
head -5 LA_Parking_Citations.csv
```

## Import data

We import the data as tibble using tidyverse:
```{r}
library("tidyverse")
system.time({lapktk <- read_csv("LA_Parking_Citations.csv")})
lapktk
```

## Fix column names

```{r}
names(lapktk) <- str_replace_all(names(lapktk), " ", "_")
lapktk
```

## Fix dates and times

```{r}
library("lubridate")
class(lapktk$Issue_Date)
lapktk <- lapktk %>% 
  separate(Issue_Date, into = c("Issue_month", "Issue_day", "Issue_year"), convert = TRUE) %>%
  separate(Issue_time, into = c("Issue_hour", "Issue_minute"), sep = -2, convert = TRUE) %>%
  mutate(Issue_DateTime = make_datetime(Issue_year, Issue_month, Issue_day, Issue_hour, Issue_minute)) %>%
  select(-(Issue_month:Issue_minute)) %>%
  select(Ticket_number, Issue_DateTime, everything())
lapktk %>% print(width = Inf)
```

Range of time period:
```{r}
range(lapktk$Issue_DateTime, na.rm = TRUE)
lapktk$Issue_DateTime %>% year() %>% table()
```

## Ticket frequency by time

```{r}
lapktk %>% 
  filter(year(Issue_DateTime) == 2015) %>%
  ggplot(aes(x = Issue_DateTime)) + 
    geom_freqpoly(binwidth = 365) + 
    labs(title = "2015 parking tickets by day")
```

```{r}
lapktk %>% 
  filter(year(Issue_DateTime) == 2015) %>%
  mutate(Issue_Month = month(Issue_DateTime, label = TRUE)) %>%  
  ggplot(aes(x = Issue_Month)) + 
    geom_bar() + 
    labs(title = "2015 Parking tickets by month")
```


```{r}
lapktk %>% 
  filter(year(Issue_DateTime) == 2015) %>%
  mutate(Issue_MDay = mday(Issue_DateTime)) %>%  
  ggplot(aes(x = Issue_MDay)) + 
    geom_bar() + 
    labs(title = "2015 Parking tickets by month day")
```


```{r}
lapktk %>% 
  filter(year(Issue_DateTime) == 2015) %>%  
  mutate(Issue_WDay = wday(Issue_DateTime, label = TRUE)) %>%  
  ggplot(aes(x = Issue_WDay)) + 
    geom_bar() + 
    labs(title = "2015 Parking tickets by weekday")
```

```{r}
lapktk %>% 
  filter(year(Issue_DateTime) == 2015) %>%
  mutate(Issue_Hour = hour(Issue_DateTime)) %>%  
  ggplot(aes(x = Issue_Hour)) + 
    geom_bar() + 
    labs(title = "2015 Parking tickets by hours")
```

## Ticket frequency by car features

```{r}
(topcm <- lapktk %>%
#  filter(year(Issue_DateTime) == 2015) %>%
  group_by(Make) %>%
  count() %>%
  arrange(desc(n)) %>%
  head(10))
```

```{r}
lapktk %>%
filter(Make %in% topcm$Make) %>%
  ggplot(aes(x = Make)) + 
  geom_bar() + 
  labs(title = "2015 Parking tickets by car make")
```

```{r}
(topcc <- lapktk %>%
#  filter(year(Issue_DateTime) == 2015) %>%
  group_by(Color) %>%
  count() %>%
  arrange(desc(n)) %>%
  head(10))
```

```{r}
lapktk %>%
filter(Color %in% topcc$Color) %>%
  ggplot(aes(x = Color)) + 
  geom_bar() + 
  labs(title = "2015 Parking tickets by car color")
```

## Create an SQLite database

Create the tibble for writing into database:
```{r}
ladb <- lapktk %>% 
  mutate(Issue_Year = as.integer(year(Issue_DateTime)), 
         Issue_Month = as.integer(month(Issue_DateTime)),  
         Issue_Day = mday(Issue_DateTime), 
         Issue_Hour = hour(Issue_DateTime), 
         Issue_Minute = minute(Issue_DateTime)) %>%
  select(Ticket_number,
         Issue_Year:Issue_Minute,
         Meter_Id:Longitude)
ladb %>% print(width = Inf)
```

Import the `ladb` table into the SQLite database:
```{r}
library("DBI")
library("RSQLite")
library("tidyverse")
if (Sys.info()[["sysname"]] == "Linux") {
  db <- dbConnect(RSQLite::SQLite(), 
                  dbname = "/home/m280-data/la_parking/LA_Parking_Citations.sqlite")
} else if (Sys.info()[["sysname"]] == "Darwin") {
  db <- dbConnect(RSQLite::SQLite(), 
                  dbname = "LA_Parking_Citations.sqlite")
}
dbWriteTable(db, "latix", ladb, overwrite = TRUE)
dbListTables(db)
dbDisconnect(db)
```

```{bash, eval = (Sys.info()[["sysname"]] == "Linux")}
ln -sf /home/m280-data/la_parking/LA_Parking_Citations.sqlite LA_Parking_Citations.sqlite
```

Size of the database file
```{bash}
ls -l LA_Parking_Citations.sqlite
```

<!-- ## Create an PostgreSQL database -->

<!-- Import the `lapktk` data set into the PostgreSQL database: -->
<!-- ```{r} -->
<!-- library("DBI") -->
<!-- library("RPostgreSQL") -->
<!-- library("tidyverse") -->
<!-- if (Sys.info()[["sysname"]] == "Linux") { -->
<!--   db <- dbConnect(RPostgreSQL::PostgreSQL(),  -->
<!--                   dbname = "laparking", -->
<!--                   host = "localhost", -->
<!--                   user = "postgres", -->
<!--                   password = "m280") -->
<!-- } else if (Sys.info()[["sysname"]] == "Darwin") { -->
<!--   db <- dbConnect(RPostgreSQL::PostgreSQL(),  -->
<!--                   dbname = "laparking", -->
<!--                   host = "localhost", -->
<!--                   user = "postgres", -->
<!--                   password = "m280") -->
<!-- } -->
<!-- dbWriteTable(db, "latix", lapktk, overwrite = TRUE) -->
<!-- dbListTables(db) -->
<!-- dbDisconnect(db) -->
<!-- ``` -->


## Read data from database

Connect to table in database:
```{r}
db <- dbConnect(RSQLite::SQLite(), dbname = "LA_Parking_Citations.sqlite")
# db <- dbConnect(odbc::odbc(), 
#                 driver = "PostgreSQL Driver",
#                 database = "./LA_Parking_Citations",
#                 host = "localhost",
#                 port = 5432)
dbListTables(db)
la_sql <- dplyr::tbl(db, "latix")
str(la_sql)
la_sql
```

## Ticket frequency by time

```{r}
# la_sql %>% 
#   filter(year(Issue_DateTime) == 2015) %>%
#   ggplot(aes(x = Issue_DateTime)) + 
#     geom_freqpoly(binwidth = 365) + 
#     labs(title = "2015 parking tickets by day")
```

## Close connection to database

```{r}
dbDisconnect(db)
```
