---
title: 'Biostat M280: An Example of Matrix Computation'
author: "Dr. Hua Zhou"
date: "January 19, 2016"
output: html_document
---

This example illustrates some common pitfalls of matrix computation using R. 

First let's create a linear regression matrix $X$ and a weight vector $w$.
```{r}
# clean up workspace
rm(list = ls())
n <- 2000
p <- 500
# slr design matrix X
X <- cbind(rep(1, n), matrix(rnorm(n * p), n, p))
# weight vector w
w <- sqrt(1:n)          # just so they're different
# weight matrix
W <- diag(w)
```

Now let's compute the matrix $X^T W^{-1} X$ in four ways.  

1. Correct, but slow. $W$ takes lots of space.
```{r}
W <- diag(w)
system.time(xwx1 <- t(X) %*% solve(W) %*% X)
xwx1[1:3, 1:3]
```

2. Correct, less slow, still takes lots of space.
```{r}
system.time(xwx2 <- t(X) %*% diag(1 / w) %*% X)
xwx2[1:3, 1:3]
```

3. Wrong, but faster. Recycles $w$ wrong.
```{r}
system.time(xwx3 <- (t(X) / w) %*% X)
xwx3[1:3, 1:3]
```

4. Correct, but looks wrong.
```{r}
system.time(xwx4 <- t(X) %*% (X / w))
xwx4[1:3, 1:3]
```
Looks different but same execution.
```{r}
system.time(crossprod(X, X / w))
```

Display system information and clean up.
```{r}
sessionInfo()
rm(list = ls())
```