---
title: "Interactive Documents With shiny"
author: "Dr. Hua Zhou"
date: "Feb 8, 2018"
subtitle: Biostat M280
output:
  ioslides_presentation:
    #smaller: true
runtime: shiny  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = FALSE)
library("tidyverse")
```

## Why Shiny?

- Better presentations of data!

- Josh Katz's maps.

- [A nice example](https://research.hackerrank.com/developer-skills/2018/).

## Inputs and outputs 

```{r, echo = FALSE}
numericInput("rows", "How many diamonds?", 5)

renderTable({
  head(diamonds, input$rows)
})
```

----

```{r, echo = FALSE}
sliderInput("bins", "Number of bins:", min = 1, max = 300, value = 30)

renderPlot({
  ggplot(data = diamonds) +
    geom_histogram(mapping = aes(x = carat), bins = input$bins)
})
```

----

- Check the Shiny cheat sheets for a list of inputs and outputs.

## Embedded shiny app

```{r, echo = FALSE, message = FALSE}
library(rmdexamples)
kmeans_cluster(iris)
```

## Building a Shiny app

- Basic parts of a shiny app: <http://shiny.rstudio.com/articles/basics.html>

- How to build a shiny app: <http://shiny.rstudio.com/articles/build.html>

- How to launch a shiny app: <http://shiny.rstudio.com/articles/running.html>

----

- Skeleton of a Shiny app:
    ```{r, echo = TRUE, eval = FALSE}
    library(shiny)
    ui <- fluidPage()
    server <- function(input, output) {}
    shinyApp(ui = ui, server = server)
    ```

- A Shiny app has 3 components:

    1. `ui`: assemble the HTML user interface for your app.

    2. `server`: a function with instructions on how to build and rebuild the R objects displayed in the UI.

    3. `shinyApp`: combines `ui` and `server` into an app.

## Try yourself

1. In RStudio, `File` -> `New File` -> `Shiny Web App...` creates a demo Shiny app using the `faithful` data set in base R.

2. Deploy your Shiny app to <shinyapps.io>.

## Get ideas for your presentation

- Browse the [Gallery](http://shiny.rstudio.com/gallery/).