---
title: "Deep Learning - Part I (Practice)"
author: "Dr. Hua Zhou"
date: "3/15/2018"
output: 
  html_document:
    toc: true
    toc_depth: 4
---

```{r setup, include=FALSE}
options(width = 120)
knitr::opts_chunk$set(echo = TRUE)
sessionInfo()
```


## Software for machine learning

Low-level: [TensorFlow](https://www.tensorflow.org), [Theano](http://deeplearning.net/software/theano/#), [CNTK](https://github.com/Microsoft/CNTK), [Caffe](http://caffe.berkeleyvision.org), [Torch](http://torch.ch), ...

High-level: [scikit-learn](http://scikit-learn.org/stable/), [Keras](https://keras.io), [PyTorch](http://pytorch.org), ...

## TensorFlow

- Developed by Google Brain team for internal Google use. Formerly DistBelief.

- Open sourced in Nov 2015.

- OS: Linux, MacOS, and Windows (since Nov 2016.

- GPU support: NVIDIA CUDA.

- TPU (tensor processing unit), built specifically for machine learning and tailored for TensorFlow.

- Mobile device deployment: TensorFlow Lite (May 2017) for Android and iOS.

- Used in variety of Google apps: speech recognition (Google assistant), Gmail (Smart Reply), search, translate, self-driving car ...

- [Machine Learning Crash Course (MLCC)](https://developers.google.com/machine-learning/crash-course/?utm_source=google-ai&utm_medium=card-image&utm_campaign=training-hub&utm_content=ml-crash-course). A 15 hour workshop available to public since March 1, 2018.

## RStudio

RStudio provides wrapper package `keras` to the Keras library. 

## MNIST - MLP

Source: <https://tensorflow.rstudio.com/keras/articles/examples/mnist_mlp.html>

In this example, we train an MLP on the [MNIST](https://en.wikipedia.org/wiki/MNIST_database) data set. Achieve testing accuracy 98.23% after 30 epochs.

- The **MNIST** database (Modified National Institute of Standards and Technology database) is a large database of handwritten digits that is commonly used for training and testing machine learning algorithms.

- 60,000 training images, 10,000 testing images. 

### Prepare data

Aquire data:
```{r}
library(keras)
mnist <- dataset_mnist()
x_train <- mnist$train$x
y_train <- mnist$train$y
x_test <- mnist$test$x
y_test <- mnist$test$y
```
Training set:
```{r}
dim(x_train)
dim(y_train)
```
Testing set:
```{r}
dim(x_test)
dim(y_test)
```
Flatten $28 \times 28$ images into $784$-vectors and scale entries to [0, 1]:
```{r}
# reshape
x_train <- array_reshape(x_train, c(nrow(x_train), 784))
x_test <- array_reshape(x_test, c(nrow(x_test), 784))
# rescale
x_train <- x_train / 255
x_test <- x_test / 255
dim(x_train)
dim(x_test)
```
Encode $y$ as binary class matrix:
```{r}
y_train <- to_categorical(y_train, 10)
y_test <- to_categorical(y_test, 10)
dim(y_train)
dim(y_test)
```

### Define the model

Define a **sequential model** (a linear stack of layers)
```{r}
model <- keras_model_sequential() 
model %>% 
  layer_dense(units = 256, activation = 'relu', input_shape = c(784)) %>% 
  layer_dropout(rate = 0.4) %>% 
  layer_dense(units = 128, activation = 'relu') %>%
  layer_dropout(rate = 0.3) %>%
  layer_dense(units = 10, activation = 'softmax')
summary(model)
```

Compile the model with appropriate loss function, optimizer, and metrics:
```{r}
model %>% compile(
  loss = 'categorical_crossentropy',
  optimizer = optimizer_rmsprop(),
  metrics = c('accuracy')
)
```

### Train and evaluation

```{r}
system.time({
history <- model %>% fit(
  x_train, y_train, 
  epochs = 30, batch_size = 128, 
  validation_split = 0.2
)
})
plot(history)
```

### Testing

Evaluate model performance on the test data:
```{r}
model %>% evaluate(x_test, y_test)
```
Generate predictions on new data:
```{r}
model %>% predict_classes(x_test) %>% head()
```

## MNIST - CNN

Let's try the convolutional neural network (CNN) on the same MNIST data. Achieve testing error 99.16% after 12 epochs.

### Prepare data

Note instead of vectorizing the images, we keep the 2D structure.
```{r}
library(keras)

# Data Preparation -----------------------------------------------------

batch_size <- 128
num_classes <- 10
epochs <- 12

# Input image dimensions
img_rows <- 28
img_cols <- 28

# The data, shuffled and split between train and test sets
mnist <- dataset_mnist()
x_train <- mnist$train$x
y_train <- mnist$train$y
x_test <- mnist$test$x
y_test <- mnist$test$y

# Redefine  dimension of train/test inputs
x_train <- array_reshape(x_train, c(nrow(x_train), img_rows, img_cols, 1))
x_test <- array_reshape(x_test, c(nrow(x_test), img_rows, img_cols, 1))
input_shape <- c(img_rows, img_cols, 1)

# Transform RGB values into [0,1] range
x_train <- x_train / 255
x_test <- x_test / 255

cat('x_train_shape:', dim(x_train), '\n')
cat(nrow(x_train), 'train samples\n')
cat(nrow(x_test), 'test samples\n')

# Convert class vectors to binary class matrices
y_train <- to_categorical(y_train, num_classes)
y_test <- to_categorical(y_test, num_classes)
```

### Define model

Define and compile model:
```{r}
model <- keras_model_sequential()
model %>%
  layer_conv_2d(filters = 32, kernel_size = c(3,3), activation = 'relu',
                input_shape = input_shape) %>% 
  layer_conv_2d(filters = 64, kernel_size = c(3,3), activation = 'relu') %>% 
  layer_max_pooling_2d(pool_size = c(2, 2)) %>% 
  layer_dropout(rate = 0.25) %>% 
  layer_flatten() %>% 
  layer_dense(units = 128, activation = 'relu') %>% 
  layer_dropout(rate = 0.5) %>% 
  layer_dense(units = num_classes, activation = 'softmax')

# Compile model
model %>% compile(
  loss = loss_categorical_crossentropy,
  optimizer = optimizer_adadelta(),
  metrics = c('accuracy')
)
```

### Train and evaluate

```{r}
system.time({
history <- model %>% fit(
  x_train, y_train,
  batch_size = batch_size,
  epochs = epochs,
  verbose = 1,
  validation_data = list(x_test, y_test)
)
})
plot(history)
```
Testing:
```
scores <- model %>% evaluate(
  x_test, y_test, verbose = 0
)

# Output metrics
cat('Test loss:', scores[[1]], '\n')
cat('Test accuracy:', scores[[2]], '\n')
```

## Generate text from Nietzsche’s writings

Source: <https://tensorflow.rstudio.com/keras/articles/examples/lstm_text_generation.html>

```{r}
library(keras)
library(readr)
library(stringr)
library(purrr)
library(tokenizers)

# Parameters --------------------------------------------------------------

maxlen <- 40

# Data Preparation --------------------------------------------------------

# Retrieve text
path <- get_file(
  'nietzsche.txt', 
  origin='https://s3.amazonaws.com/text-datasets/nietzsche.txt'
  )

# Load, collapse, and tokenize text
text <- read_lines(path) %>%
  str_to_lower() %>%
  str_c(collapse = "\n") %>%
  tokenize_characters(strip_non_alphanum = FALSE, simplify = TRUE)

print(sprintf("corpus length: %d", length(text)))

chars <- text %>%
  unique() %>%
  sort()

print(sprintf("total chars: %d", length(chars)))  

# Cut the text in semi-redundant sequences of maxlen characters
dataset <- map(
  seq(1, length(text) - maxlen - 1, by = 3), 
  ~list(sentece = text[.x:(.x + maxlen - 1)], next_char = text[.x + maxlen])
  )

dataset <- transpose(dataset)

# Vectorization
X <- array(0, dim = c(length(dataset$sentece), maxlen, length(chars)))
y <- array(0, dim = c(length(dataset$sentece), length(chars)))

for(i in 1:length(dataset$sentece)){
  
  X[i,,] <- sapply(chars, function(x){
    as.integer(x == dataset$sentece[[i]])
  })
  
  y[i,] <- as.integer(chars == dataset$next_char[[i]])
  
}

# Model Definition --------------------------------------------------------

model <- keras_model_sequential()

model %>%
  layer_lstm(128, input_shape = c(maxlen, length(chars))) %>%
  layer_dense(length(chars)) %>%
  layer_activation("softmax")

optimizer <- optimizer_rmsprop(lr = 0.01)

model %>% compile(
  loss = "categorical_crossentropy", 
  optimizer = optimizer
)

# Training & Results ----------------------------------------------------

sample_mod <- function(preds, temperature = 1){
  preds <- log(preds)/temperature
  exp_preds <- exp(preds)
  preds <- exp_preds/sum(exp(preds))
  
  rmultinom(1, 1, preds) %>% 
    as.integer() %>%
    which.max()
}

for(iteration in 1:60){
  
  cat(sprintf("iteration: %02d ---------------\n\n", iteration))
  
  model %>% fit(
    X, y,
    batch_size = 128,
    epochs = 1
  )
  
  for(diversity in c(0.2, 0.5, 1, 1.2)){
    
    cat(sprintf("diversity: %f ---------------\n\n", diversity))
    
    start_index <- sample(1:(length(text) - maxlen), size = 1)
    sentence <- text[start_index:(start_index + maxlen - 1)]
    generated <- ""
    
    for(i in 1:400){
      
      x <- sapply(chars, function(x){
        as.integer(x == sentence)
      })
      x <- array_reshape(x, c(1, dim(x)))
      
      preds <- predict(model, x)
      next_index <- sample_mod(preds, diversity)
      next_char <- chars[next_index]
      
      generated <- str_c(generated, next_char, collapse = "")
      sentence <- c(sentence[-1], next_char)
      
    }
    
    cat(generated)
    cat("\n\n")
    
  }
}
```